Learning Heuristics of SAT by Graph Neural Networks

### 2018-12-14 report/paper.tex
update

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  In this document, we summarize the discussion and provide some formal description of the proposed model.
\end{abstract}

@@

\section{Problem Statement}

In this paper, we report the progress and result on the project that aims to learn branching heuristics 
for SAT problem based on DPLL algorithm. The DPLL algorithm \cite{Davis:1962:MPT:368273.368557} and its 
variants are widely used in many modern SAT solvers. 
The core ideas behind DPLL are searching and unit propagation. 
The searching involves choosing a variable and assigning it to either $\top$ or $\bot$, 
which is the part of our concern in this paper.

The DPLL algorithm works on Conjunction Normal Form. A CNF formula is a conjunction of clauses. A clause is a disjunction of literals. A literal is either a variable (symbol) or a negation of a variable. Formulas in other forms can be translated to CNF \cite{tseitin1983complexity}. For example, $(x_1 \vee x_2) \wedge (\neg x_2 \vee x_3)$ is a valid CNF formula.

@@

Particularly, we would like to learn the branching heuristics for \texttt{f.pickVariable}.
In this paper, we use graph neural network to encode a CNF formula and use reinforcement learning 
to learn the best branching heuristics.
We use the policy gradient method for the reinforcement learning part.

\section{Proposed Model}

@@

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
%\begin{abstract}
%\end{abstract}

@@

\section{Introduction}

In this paper, we report the progress and preliminary result on the CS590-DLS course project that aims 
to learn branching heuristics for the Boolean Satisfiability (SAT) problem based on DPLL algorithm.

\paragraph{Background}

Given a formula consists of variables over booleans and propositional logic connectives, the satisfiability
problem asks an assignment, i.e., a mapping from variables to boolean values, that makes the formula be true. 
For example; $(P \wedge Q) \to (\neg Q \vee R)$ is a well-formed formula and can be satisfied by assigning
$P$ to be $\top$; $P \wedge \neg P$ is another well-formed formula but can not be satisfied, written as UNSAT.

A traditional approach to solve SAT is DPLL algorithm \cite{Davis:1962:MPT:368273.368557}. 
Although in the worst case, DPLL is still an exponential algorithm due the nature of the problem,
but on many instances DPLL is able to find a solution or prove UNSAT quickly.
Thus DPLL and its variants are widely used in many modern SAT solvers such as miniSAT.

The DPLL algorithm works on Conjunction Normal Form (CNF). 
A CNF formula is a conjunction of clauses. A clause is a disjunction of literals. 
A literal is either a variable (symbol) or a negation of a variable. 
Formulas in other forms can be translated to CNF \cite{tseitin1983complexity}. 
For example, $(x_1 \vee x_2) \wedge (\neg x_2 \vee x_3)$ is a well-formed formula in CNF.

With the CNF format introduced, now we can briefly describe the essential ideas behind DPLL: 
unit propagation, searching and backtracking.
If a clause contains only one literal, then it is a unit clause. Since the problem is given 
in CNF, this cluase is forced to be satisfied, thus we must assign this literal to be true. 
The unit propagation then calculate the consequential formula by making this assignment --
some clauses can be complete eliminated.
If the current formula does not contain an unit clause, the algorithm then chooses a branching
variable by predefined heuristics and assign either $\top$ or $\bot$ to this variable,
and then propagate this decision and recursively call the decision procedure.
At some point, we may reach a dead end that the current assignment shows an UNSAT result, 
then we need to recover to some previous decision we made and try another assignment.

@@

Several variants based on this algorithm make it more efficient: 1) one can introduce
pure literal elimination, that is a variable appears with only one polarity can be 
assigned to a value that makes its containing clauses satisfies but without triggering
any conflict. 2) When we reach one dead end, we can perform a conflict analysis which 
produces a lemma that must be satisfied (known as conflict-driven clauses learning, CDCL
\cite{DBLP:series/faia/SilvaLM09}). 
The lemma will be added back to the formula, and the backtrack can be done in a non-chronological fashion.
In this paper, we constrain ourself to the simplest version of DPLL and focus on
learning how to choose the branching variable.

\paragraph{Problem Statement}

The fundamental problem we are interested in is

Particularly, we would like to learn the branching heuristics for \texttt{f.pickVariable}.
In this paper, we propose to use recent advances in graph neural network to 
encode a CNF formula and use reinforcement learning to learn the best branching heuristics.
We use the policy gradient method for the reinforcement learning part.

\section{Model}

In this section, we briefly review the related work on graph representations of SAT
and then describe the model used in this paper.

@@

\section{Implementation}

To make the DPLL algorithm integrates well with the ML/RL part, we implement a small-step version of 
DPLL, but without changing the decision procedure. This can be obtained by applying 
defunctionalization \cite{DBLP:conf/ppdp/DanvyN01} to the recursive DPLL implementation, which exposed a continuation
component that saves the decision history and the current state. Another similar presentation
is \citeauthor{DBLP:conf/lpar/NieuwenhuisOT04}'s abstract DPLL.
This allows us easily inspect the current state and explicitly handle backtrack.

\section{Experiment}

\section{Related Work}

As artifical neural networks greatly succeeded in vision and natural language tasks in last ten years,
many works that combine deep learning and logical reasoning also has been emerged recently.

\section{Conclusion}

\textbf{Acknowledgement} The author would like to thank Fei Wang for the inspiring discussion and 
help of the machine learning part.
